---
title: "Assignment 3"
author: "Ryan Trieu (dtrieu99@uw.edu)"
date: "03/01/2022"
output: html_document
---
**Affiliation**: "INFO-201: Technical Foundations of Informatics - The Information School - University of Washington"

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(knitr)
source("../docs/analysis.R")
```


# 1.0 Introduction 

According to a study from the National Council on Crime and Delinquency in 2006[^1], the United States has come to rely on jail as a reaction to all sorts of crime in the last 30 years. Minor infractions of parole or probation frequently result in a return to prison. This has resulted in a massive jail system in the United States. The research also mentions interesting facts about incarceration in the United States:
- The United States is home to the world's largest prison population.
- The incarceration rate in the United States is four times that of the rest of the world.
- Some US states imprison up to six times the number of individuals as countries with similar populations.
- The United States has the highest rate of female incarceration in the world.
- Incarceration rates are not taken into account when calculating crime rates.
With such a high population of incarcerations, there are bound to be different variables that draw attentions of scholars, activists, and the public in general. In this report, I will discuss 

# 2.0 Summary Information
County with biggest jail population by race in 2018
```{r, echo = FALSE, warning=FALSE, message=FALSE}
kable(biggest_jail_pop_race_2018)
```

County with biggest jail admission by race all time
```{r, echo = FALSE}
kable(biggest_prison_adm_race_2018)
``` 

Total prison admission by race all time
```{r, echo = FALSE} 
kable(total_prison_adm_race)  
```


Total jail population by race all time
```{r, echo = FALSE} 
kable(total_jail_pop_race)
```

Top 10 counties with the highest total jail from ice in 2018
```{r, echo = FALSE} 
kable(biggest_jail_ice)
```

Top 5 counties in Washington State with highest jail population in 2018
```{r, echo = FALSE}
kable(top_5_jail_pop_data_WA_2018)
```

# 3.0 Trends over time chart
Trend of black jail pop vs white jail pop in King County from 2014-2018
```{r, echo = FALSE} 
trend_chart
```

# 4.0 Variable comparison chart
Rated capacity vs Total Jail Admissions Count
```{r, echo = FALSE} 
variable_chart
```

# 5.0 Map
Map of top 5 states with biggest Jail Population Count Held for ICE
```{r, echo = FALSE} 
map
```

## References
[^1]: [Hartney, Christopher. “US Rates of Incarceration: A Global Perspective.” (2006).](https://www.semanticscholar.org/paper/US-Rates-of-Incarceration%3A-A-Global-Perspective-Hartney/cbcc79e99c3ef29a421d2d306db27bea141e5a0d)